Homicide Increases in the COVID-19 Era

### 2025-02-20 Draft1.qmd
added my stuff

>>>>

# Plot 1 - Examining Crime Density from 2010–2019 and 2020-Present

```{r}

par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))

smoothScatter(
  x    = crime_la_2010_2019$LON,
  y    = crime_la_2010_2019$LAT,
  xlim = c(-119, -118),
  ylim = c(33.5, 34.5),
  nbin = 300,
  xlab = "Longitude",
  ylab = "Latitude",
  main = "Crime Density (2010–2019)",
  colramp = colorRampPalette(c("white", "skyblue", "blue", "darkblue"))
)

smoothScatter(
  x    = crime_la_2020_present$LON,
  y    = crime_la_2020_present$LAT,
  xlim = c(-119, -118),
  ylim = c(33.5, 34.5),
  nbin = 300,
  xlab = "Longitude",
  ylab = "Latitude",
  main = "Crime Density (2020–Present)",
  colramp = colorRampPalette(c("white", "pink", "red", "darkred"))
)

par(mfrow = c(1, 1))

```
Both time periods show major clusters of reported crimes in central LA. The 2020–present data appears slightly more dispersed to the north, but overall patterns remain similar looking.

# Plot 2 - Examining Number Of Crimes Over Time

```{r}

crime_la <- rbind(crime_la_2010_2019, crime_la_2020_present)

crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
                                  format = "%m/%d/%Y %I:%M:%S %p", 
                                  tz = "America/Los_Angeles")

crime_la$YearWeek <- format(crime_la$DateParsed, "%Y-%U")

weekly_counts <- aggregate(DR_NO ~ YearWeek, data = crime_la, FUN = length)
names(weekly_counts)[2] <- "Crimes"

weekly_counts <- weekly_counts[order(weekly_counts$YearWeek), ]

weekly_counts$Index <- seq_len(nrow(weekly_counts))

plot(
  weekly_counts$Index,
  weekly_counts$Crimes,
  type = "l",
  col  = "blue",
  lwd  = 2,
  xlab = "Weekly Index",
  ylab = "Number of Crimes",
  main = "Weekly Crime Counts in Los Angeles (2010–Present)"
)

abline(h = pretty(weekly_counts$Crimes), v = pretty(weekly_counts$Index), 
       col = "gray", lty = "dotted")

```
The time series shows stable weekly crime counts with occasional large spikes. These surges indicate certain weeks where crime reporting jumped significantly, there is a recent interesting surge that seems to mean-revert in 2024.

# Using Regression to Predict ____

```{r}

weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])

ar1_trend_regression <- lm(Crimes ~ Index + CrimesLag, data = weekly_counts, na.action = na.omit)

summary(ar1_trend_regression)

```

# Interpretation

The lagged crime coefficient (0.81949) suggests that a high crime week is often followed by another high crime week. Since the trend term is not significant, there is no clear drift in weekly crime over time.
